---
title: "Exponential Distribution Simulation and Comparison"
output: html_document
---

## Overview
Investigate the exponential distribution in R and compare it with the Central Limit Theorem.The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials. You should:

1. Show the sample mean and compare it to the theoretical mean of the distribution.

2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

3. Show that the distribution is approximately normal.



```{r setup}
# Setup Code
library(ggplot2)

# Variables
sims <- 1000
lambda <- 0.2
n <- 40

set.seed(3737)


```

# Simulation
Create the matrix of 1000 rows. 
We will use 40 columns as the samples of the random sim.  
```{r}
# Since a matrix fills in a cascade, this does the job of generating the 
# sims and the samples.
sim_matrix <- matrix(rexp(sims * n, rate=lambda), sims, n)

# Means of the random samples.
sim_mean <- rowMeans(sim_matrix)
```

Let's plot the simulation means:
```{r}
hist(sim_mean)
```


