---
title: "Exponential Distribution Simulation and Comparison"
output: html_document
---

## Overview
Investigate the exponential distribution in R and compare it with the Central Limit Theorem.The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials. You should:

1. Show the sample mean and compare it to the theoretical mean of the distribution.

2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

3. Show that the distribution is approximately normal.



```{r setup}
# Setup Code
library(ggplot2)

# Variables
sims <- 1000
lambda <- 0.2
n <- 40

set.seed(3737)


```

# Simulation
Create the matrix of 1000 rows. 
We will use 40 columns as the samples of the random sim.  
```{r}
# Since a matrix fills in a cascade, this does the job of generating the 
# sims and the samples.
sim_matrix <- matrix(rexp(sims * n, rate=lambda), sims, n)

# Means of the random samples.
sample_means <- rowMeans(sim_matrix)
```


Let's plot the simulation means:
```{r}
#sim_mean_df <- data.frame(sim_mean)
#g <- ggplot(sim_mean_df, aes(y = ..ndensity.., x=sim_mean)) + geom_histogram()
#print(g)

#TODO -- ggplot vs hist return different counts.  figure out why.

hist(sample_means, col="blue")
```

# Sample Mean (Center) vs Theoretical Mean (Center)
```{r}
theoretical_mean <- 1/lambda
sample_mean <- mean(sample_means)
```

We can see that the theoretical mean is **`r theoretical_mean`** and the mean of the samples is **`r sample_mean`**.  The difference between the two is minimal - only **`r abs(sample_mean - theoretical_mean)`.**  TODO -- is this an unbiased estimator? Or can this be considered an estimator <> estimand?

# Sample Variance vs Theoretical Variance
```{r}
sample_var <- var(sample_means)
theoretical_var <- (1/lambda) ^ 2/n
```

We can see that the theorical variance is **`r theoretical_var `** and the variance of sample means is **`r sample_var`**.  The difference between the two is minimal - only **`r abs(sample_var - theoretical_var)`.**  TODO -- is this an unbiased estimator? Or can this be considered an estimator <> estimand?

# Distribution: 
How can we tell the distribution of the sample means is approximately normal? 

Let's simulate and plot a random normal distribution, and show a distribution of our adjusted sample means based on the CLT formula of

***Adjusted from slides -- check if right***

$$\frac{\bar X_n - \mu}{\sigma / \sqrt{n}}=
\frac{\sqrt n (\bar X_n - \mu)}{\sigma}
= \frac{\mbox{Estimate} - \mbox{Mean of population }}{\mbox{Std. Err. of estimate}}$$ 

next to it.

```{r fig.width=4, fig.height=4, echo=FALSE, out.extra='style="float:left"'}
normal_sim <- data.frame(x = rnorm(1000, 0, 1))
hist(normal_sim$x, col="red", main="Simulated Random Normal Distribution")
```

```{r fig.width=4, fig.height=4, echo=FALSE, out.extra='style="float:left"'}
adjusted_sample_means <- 
    data.frame(x = (sample_means - (1/lambda))/theoretical_var/sqrt(40))
hist(adjusted_sample_means$x, col="blue", main="Adjusted Sample Means Per CLT")
```

We can see from the images above that both a randomly generated normal distribution and our adjusted sample means both center around 0, and follow the same pattern.  
